To Airbnb?: A Question of Revenue

### 2018-05-31 papers/full_paper.Rmd
Title Change and Final updates for anonymous submission

---
title: "To Airbnb?: A Question of Returns"
subtitle: "Andy Krause and Gideon Aschwanden"
author: "Greenfield Advisors and University of Melbourne"
date: "May 5th, 2018"
output:
  pdf_document: default
  html_document: default
header-includes: 
   - \usepackage{setspace}\doublespacing
   - \usepackage{caption}
   - \captionsetup[table]{skip=10pt}
geometry: margin=3.3cm
bibliography: abb.bib
editor_options: 
  chunk_output_type: console
---

@@

@@

With the rise of Airbnb and other self-hosting accommodation platforms^[Flipkey.com, VRBO.com, Homeaway.com, Tripping.com, Stayz.com.au as examples.], it is easier than ever for property owners and investors to consider short-term rental (STR) as a potential leasing strategy for a revenue generating residential property. Should they though? Or rather, is it more profitable to do so when compared to traditional long-term rental (LTR)? And further, are certain types of properties in specific locations more likely to show a preference for a short-term leasing strategy over long-term one.

@@

We adopt a scenario-based approach that compares observed annual revenues for short-term rental (STR) properties with an estimate of their expected long-term rental revenues. Our results show that the feasibility of a greater return through a short-term rental -- what we have termed ``Short-Term Preference'' -- is dependent on property type, product, location and host policies.  In short, the answer to whether or not a owner or investor should opt for a short-term leasing strategy: it depends.

@@

Existing research into short-term rental markets is sparse. Within the existing work, the focus remains centered on rural and/or amenity-based STRs such as beach-front and ski-resort access properties.  This focus resulted from the fact that, until a few years ago, the STR market itself operated almost wholly within these market segments. The shift to small, urban dwellings within the STR market is relatively new and explains the dearth of past work in this direction. With the arrival of Airbnb, and the controversies surrounding it, the topic is creating interest both in real estate and economics as well as in a wide variety of fields as evidenced by a growth in recent publications and on-going working papers.

@@

Focusing specifically on the short-term rental market, @cassidy2007tourist suggest that within Australia's Gold Coast, the STR industry lacks sophistication and efficiency.  Most notably, they find that most nightly or weekly rates are set by resident unit managers (RUMs) and are often under-priced as there is little accountability to owners.  The authors also note that very little research has been put toward understanding this growing market segment.

@@

Though issues of second homes have been primarily treated as a rural phenomenom for decades  [@gallent2007second], a wave of ``new urban tourism.'' Fuller and Michel [-@fuller2014stop] has re-focused some of the problems of short-term rentals on large, urban environments. With this move to urban areas, issues such as nuisance, property rights and regulation have naturally arisen.  As a result, the short-term rental market has garnered significant interest in the public policy and legal sphere as of late. As many second homes are not permitted for or regulated as commercial entities, STRs can have serious effects on immediate neighbors in the way of externalies such as noise and congestion [@frost2006commentary].  While strict regulation is often proposed, doing so may constititute regulatory takings [@jefferson2015can].  Other, rights-based, regulatory approaches have been suggested [@miller2014transferable].  Many of the key legal decisions regarding short-term rentals, particulary through Airbnb, are still in progress and much remains to be settled in this area in terms of the regulatory and legal environment in which STRs are allowed to operate.

@@

Australia implements a particular tax policy, often termed `negative gearing', that allows for income tax deduction to real estate investors who experience greater costs than revenues on investment properties.^[See @fane2004negative for a complete explanation of negative gearing.] To date, this policy only applies to long-term rental properties.  While this policy can influence the net revenue (post tax) of a property, the negative gearing policy requires knowledge of the investor's personal income situation.  As we do not have access to this information in the marketplace, we will not be considering any effects of this policy in our analysis here.

One benefit of a short-term rental may be the ability to adjust the nightly rate over the course of the year to reflect changing demand and market trends. As future market trends are unknown at the time of purchase and, within the span of 12 months, the decision of long- vs short-term leasing strategy is asymetrically irreversable (can go from short to long, but not long to short), we make our comparisons using a fixed nightly (short-term) rate to respresent the most likely state of knowledge for our hypothetical investor on September 1, 2015. Both long-term rents and short-term rates and occupancy structures will be based on the market conditions as of this date.

@@

Janitorial costs, $J$, cover the costs to 'turn-over' the property between occupants.  For a short-term leasing strategy, these costs can be experienced every day or two, depending on the minimum length of stay and the total number of occupancies over the course of the year. Our data for short-term rentals comes from the Airbnb platform.  In addition to the nightly rate that an occupant pays to the owner there is also a separate 'cleaning fee' that is attached, as is a booking fee to Airbnb.^[This fee could be considered analogous to the leasing fees $L$ that long-term rentals pay.] Some hosts add additional 'linen' fees to cover costs of laundering.  While it may be advantageous for hosts in competitive markets to underquote their cleaning fees to keep occupancies up, we have no evidence to suggest that owners are exhibiting such behavior. As a result of janitorial fees being a separate line item in the STR data and directly paid by the occupants they can also be ignored in the comparative analysis below.

@@

\begin{equation}
U = (4.5 * B * 1.7) * O
\end{equation}

@@

Unfortunately, there are no common identifiers between the two datasets as the Airbnb data does not include address. Both sets of data do include latitude and longitude, but in a dense urban area with many multiple-family dwellings it is impossible to match apartments based on two-dimensional coordinates. As a result of this, we are unable to directly match observations from the two datasets. However, the number of properties that rented long-term for the 2014 to 2015 period and then switched to Airbnb for the 2015 to 2016 period are likely very small and this inability to directly match properties does not greatly impact our modeling strategy.

@@

Our data preparation primarily involved the filtering out of data errors, outlier values and other observations than contain data that do not fit our research questions. Additionally, as our data are from two different sources, a number of fields require standardization so that comparisons between the two leasing strategy options -- long-term rental (LTR) vs short-term rental (STR) -- can be made.

We begin by filtering observations by time. Our hypothetical leasing strategy decision is that of an investor who purchases a property on September 1, 2015 and must decide on traditional long-term rental versus an Airbnb-type, short-term rental approach to generating income. To determine the likely LTR rate for any given property, we will use the observed rental transactions from September 1, 2014 to August 31, 2015. Taking the Airbnb or STR approach allows the owner to change rates over time and is subjected to daily changes in the market (supply and demand factors). In this simulated example we will use Airbnb data from September 1, 2015 to August 31, 2016 to represent the actual Airbnb market conditions over the time period in question.

@@

After removing property level observations that are missing price data, we find that most long-term weekly rentals are greater than \$200 and less than \$1,000 per week while most median nightly rates are greater than \$50 and less than \$500 per night. We filter the property level data accordingly. We also filter the short-term daily data to include only those observations with nightly rates between \$50 and \$500.

@@

To better visualize the relative frequency of the submarkets in the two tenure types we have broken down the data by submarket by property type by tenure type. A number of interesting observations are illustrated here (Figure 4). First, the vast majority of Airbnb units are apartments in city-core or city locations. As expected and as shown by the map, the long-term properties are more evenly spread across the metro region than the short-term properties.  Long-term houses are more peripheral than apartments, a result of urban development patterns and land economics. Beach properties are, relatively speaking, over-represented in the short-term market, especially in apartments, while suburban and rural apartments, somewhat common in long-term markets, are nearly absent in the short-term market. Overall, the suburban and rural Airbnb markets are quite thin, later results focusing on these markets will need to be examined in light of this small sample size.

@@

<!-- ### Summary of Data Cleaning -->

<!-- We have applied ten separate filters to our two datasets. In the name of transparency and to highlight the bounds of our research question it is useful to review these filters (Table 2). There are more than 20,000 properties that are or have been listed on the Airbnb portal in Melbourne during the 2014 to 2016 period.  Filtering those not listed during the study period removes nearly 1/3 of all properties.  Further removing by property type and listing type (not entire home) elimates another 50% of the remaining properties.  From these 7,200 properties structural characteristics filters remove another 480 from consideration.  Over 1,000 of the remaining properties were never been booked and/or had outlying nightly rates and were removed.  Finally, 14 properties had either bad location data or missing information host policy. From over 20,000, we have 5,788 Airbnb properties meet the requirements of our research question. -->

<!-- For the long-term data, over half of the observations fall outside of our considered time frame, reducing the sample size down to around 130,000 from 290,000.  Further consideration regarding property types, structural characteristics, location, etc. remove another 12,000 observations, leaving 120,092 in the final dataset. -->

@@

First, we must deal with those properties that were not listed for the entire 12-month period.  Our data cleaning eliminated all properties that were listed for less than six months, however, it would be incorrect to compare the revenue of a property listed on Airbnb for seven months with the hypothetical annual revenue from a long-term rental.

@@

1. Actual ($REV_o$)}: Observed bookings multiplied by median nightly rate
2. Likely ($REV_l$)}: Observed bookings extrapolated to a full year representing the likely actual revenue if the property had been listed all year long,^[Properties listed in summer, but not winter may be over imputed and vice versa, however, we expect the biases to cancel out over the entire sample.] based on observed occupancy rate multiplied by median nightly rate
3. Potential ($REV_p$)}: Observed occupancy rate of non-blocked days multiplied by 366 multiplied by median nightly rate

As the 'Actual' calculation systematically under counts the revenue from properties that were not on the Airbnb portal for the entire year we disregard it going forward and consider the 'Likely' revenue metric to be a better measure of actual market performance. We chart the distribution of the Likely ($REV_l$) and Potential ($REV_p$) short-term revenues against a distribution of the observed long-term revenues for the long-term properties (Figure 5). Revenues from the long-term rentals are far more consistent, with the vast majority totaling between \$12,000 to \$30,000 per year.  Likely short-term revenues range from neglible to highs of greater than \$80,000, with many of the properties showing very limited (less than \$10,000) revenue. The range of the distribution for the Potential short-term revenue is similar, however, the distribution itself is much flatter, with many more properties showing potential incomes in the \$20,000 to \$40,000 range.

While this simple analysis does not control for location and product differences, it does highlight the fact that some Airbnb owners are likely making greater revenues than in long-term rentals, but that most are likely not.  When looking at potential revenues, however, a great deal more may have the potential to produce additional revenue.  The high number of very low yield Airbnb properties in the Likely condition suggests that some owners may not be focused on maximizing revenue, but instead prefer the flexibility of short-term leasing.^[Do note that we have not considered any overall market demand issues when analyzing potential revenue.]

![Observations by Submarket](C:/Code/research/airbnbMelbourne/images/rev_dens.png)

@@

Realizing that the imputed long-term rental values may be more accurate when estimated at smaller scales, we impute long-term rental rates at submarket level, specifically by type and geographic area. In other words, we separately estimate the LTR model in Equation 10 for Apartments in the city core, Apartments with  in the beach submarket, etc. for all combinations of type and geographic area. We then combine the results from all estimates back into a single dataset that contains all of the short-term rentals.

@@

where $GPR$ is the total number of guests allowed per bedroom^[Hosts set the total number of guests.  We have normalized this by bedroom to avoid co-linearity in our model with the bedroom variables.], $MinStay$ is the minimum number of nights that a stay may have and $CancelPolicy$ is the difficulty with which a booking may be cancelled.  We have grouped the cancellation policies into three groups, Flexible (24 hour notice), Moderate (5 day notice) and Strict (50\% refund, 7, 30 or 60 day notice required).

The results of all three models are shown in Table 4. For $PType$ Apartments are the reference category,  for $BedBath$, 1 Bed / 1Bath is the reference category, for $Submrkt$ the City-Core is the reference category and for $CancelPolicy$ Moderate is the reference category.  As the additional variables are added, the model fit diagnostics steadily increase, suggesting that the increased complexity of the models adds to the quality of the fit.

In Model 1, we see that when controlling for structural characteristics, Houses have a significantly lower probability of being more profitable in short-term leasing than apartments. Additionally, 1 bed / 1 bath units show greater likelihood of short-term preference than 2 bed / 1 bath and 2 bed / 2 bath combinations. This could be due to the fact that 2 bed units may be very attractive in the long-term market as they are a useful product for two, non-family roomates to share. Larger units, 3 bedrooms plus, generally show no statistical difference in preference porbability to the 1 bed / 1 bath reference unit, with the exception of the 4 bed / 2 bath properties.

Adding the submarket variables (Model 2) shows that all areas have a much lower probability of short-term prefernce when compared to the City Core (the reference category) when controlling for structural characteristics.  The difference between the Rural and City-Core is not significant, however. The coefficients for the structural characteristics change little from Model 1 to Model 2, speaking to the robustness of those estimates. The coefficient for House properties does decrease significantly in magnitude when adding the geographic submarkets.  This is likely due to the fact that there are sturctural differences between the geographic submarkets.  Houses, for example, are concentrated more heavily in City and Suburban areas than in the City-Core and some of the impact of House is being mediated by the geographic submarkets.

Finally, in Model 3 we add in host behavior or policies. Allowing a greater number of guests per bedroom results in a higher probability of short-term preference.  This variable may be operating as proxy for additional sleeping spaces which allow for a higher nightly rate, ceteris paribus, to be charged.  The greater the minimum stay, the lower the probability of short-term preference.  Longer minimum stays may equate to lower occupancy rates as one or two night travelers are exempted from the property.  Owners with longer minimum stays may be looking to minimize costs such as cleaning and administration.  Flexible cancellation policies result in a lower probability of short-term preference, while a Strict policy has a positive relationship with short-term preference.  Strict policies do help maintain revenues, but may also scare off some potential guests.  Strict cancellation may be a luxury for the best and highly desired properties on the market and, within our model, may be acting as a proxy for some unobserved measure of quality or demand while a flexible cancellation may be signalling the opposite.

@@

  # XTable
  x.full <- xtable(full.table,
                  caption='Logistic Modeling Results\n')
  align(x.full) <- 'lllll'

@@

The above uses the Likely^[Extrapolated out in cases of Airbnb properties on the market for less than a year] revenue in the comparisons with the imputed long-term revenue.  Doing so may underestimate the possible revenue -- what may matter most to the hypothetical buyer in our scenario -- due to not factoring in the impact of blocked days.  If we, instead, assume that blocked days could be leased out on the short-term market at the same overall occupancy rate as the unblocked (observed) days we can produce a potential revenue figure. Note that it is possible (likely?) that the large oversupply of available units in the market that would occur if no properties blocked their units would decrease both prices and occupancy rates.  We do not attempt to correct for this likelihood, rather we offer these figures as high end estimates.

Using the potential revenue, 53.3% of all properties are short-term preferable.  Breaking that down by type and by geographic submarket (Table 3) shows similar findings, in terms of relative comparativeness, to the Likely revenue analysis above.  Apartments have a higher short-term preference than do Houses and City-Core properties dominate the other geographic regions.  One notable difference here is that the Rural properties perform rather poorly under a potential revenue scenario, relatively speaking, suggesting that they may be operating near their peak in the current condition.

@@

 px.full <- xtable(pfull.table,
                   caption='Logistic Modeling Results\n')
 align(px.full) <- 'lllll'

@@

Overall, the choice of revenue type, Likely or Potential, does affect some of the conclusions to be derived from the results above.  As expected, under the Potential case there are more properties (53\% vs 28\%) that exhibit short-term preference.  Under the scenario, the possibilities of increasing revenue by moving from long- to short-term leasing appear much better.  Again, however, we have not attempted to measure or estimate how this hypothetical increase of properties in the market (due to the elimation of blocked days) would impact overall market dynamics.  Suffice it to say that it would like drive down occupancies and maybe price, therefore leading to a decrease in short-term preference likelihood.

@@

If these are the dominant two forms of hosts in the market, then a distribution of properties by the percent of time which they are blocked should be highly bi-modal with profit seekers on the left and opportunistic sharers on the right. The left hand panel of \autoref{hostblock} shows that this breakdown holds somewhat, with peaks of hosts at each end, but that there are also a considerable number of hosts who block their properties some (25\% to 75\%) of the time.

There is also a third potential host type, that of the host who utilizies multiple platforms to list their short-term property, of which Airbnb may just be one. A potential method to identify these hosts is to look at the number of individual periods of blocked days.  A host using multiple platforms will likely have many small periods of blocked days on Airbnb as they will block out those days in which reservations are made on other platforms. We test to see if this is a meaningful metric by plotting the total count of blocked periods versus the percentage of blocked days in the right panel of \autoref{hostblock}. We do find many hosts with a large number of blocked periods, but, unfortunately, no true pattern emerges.

@@

Attempts to divide the hosts up with unsupervised^[As there are no official host types we could not run more reliable supervised models.] clustering algorithms -- including the use of more variables -- failed to produce robust categories of hosts. As a more subjective, but implementable, approach we have divided the hosts up into four categories based on the original two variables considered: block percentage and number of blocked periods. Profit seekers are those host who have blocked percentages less than 25\% of the time.  Opportunistic Sharers have blocked their properties at least 75\% of the time.  Within those two extremes, we consider users who have at least 12 blocked periods during a year to be 'Multi-Platform Users' and the remaining set to be 'Unknown'. Profit Seekers make up 43\% of the hosts, with Opportunistic Sharers capturing 18\% and Multi-Platform Users 10\%.  The remainder 'Unknown' category contains just over 28\% of all properties. See \autoref{hostcat} and \autoref{htc} for a breakdown of these group by the two variables.

@@

Looking at the differences in short-term preference under both revenue cases, we see large variation in short-term preference between host types in the Likely case and a much smaller variation in the Potential case, as expected. There is little change for the 'Profit Seekers' as most of these hosts are already operating with few blocked days.  For the other three sets, considerable gains could be had with a more aggresive leasing approach. As the potential figures are somewhat similar here, we will only examine the differences for host types in the Likely number going forward.

@@

Table 6 compares logistic models results from three of the host types.  The Opportunistic Sharers show no instances of short-term preference and therefore could not be estimated.  The left hand column, that of the Profit-Seekers, is the most representative of a sample of property owners that are seeking to maximize revenue.  Overall, the coefficients from this sub-sample match up well to the results from the entire sample shown in Table 5.  In short, the conclusions are the same: small apartments in the central city with short minimal stays, inflexible cancelation policies and a high guest limit offer the highest probability of short-term preference.  The models for the Multi-Platform users (middle panel) and the Unknowns (right) are not as clear as the Profit-Seekers but do suggest overall similar dynamics.

@@

 xh.full <- xtable(hfull.table,
                  caption='Logistic Model Results of Potential Short-Term Preference\n')
 align(xh.full) <- 'lllll'

@@

In short, about 28\% of current Airbnb properties are outperforming their potential in the long-term market.  If owners all operated as profit maximizers this figure could be 54\% or higher given that the overall demand for short-term rental could match the assumed supply increase in such a case.

>>>>

---
title: "To Airbnb?: A Question of Revenue"
subtitle: "Andy Krause and Gideon Aschwanden"
author: "Greenfield Advisors and University of Melbourne"
date: "May 5th, 2018"
output:
  pdf_document: default
  html_document: default
header-includes: 
   - \usepackage{setspace}\doublespacing
   - \usepackage{caption}
   - \captionsetup[table]{skip=10pt}
geometry: margin=3.3cm
bibliography: abb.bib
editor_options: 
  chunk_output_type: console
---

@@

```{r options_functions, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE, results='hide'}

@@

With the rise of Airbnb and other self-hosting accommodation platforms^[Flipkey.com, VRBO.com, Homeaway.com, Tripping.com, Stayz.com.au as examples.], it is easier than ever for property owners and investors to consider short-term rental (STR) as a potential leasing strategy for a revenue generating residential property. Should they though? Or rather, is it more profitable to do so when compared to traditional long-term rental (LTR)? And further, are certain types of properties in specific locations more likely to show a preference for a short-term leasing strategy over long-term one?

@@

We adopt a scenario-based approach that compares observed annual revenues for short-term rental (STR) properties with an estimate of their expected long-term rental revenues. Our results show that the feasibility of a greater return through a short-term rental -- what we have termed ``Short-Term Preference'' -- is dependent on property type, product, location and host policies.  In short, the answer to whether or not a owner or investor should opt for a short-term leasing strategy is ''it depends''.

@@

Existing research into short-term rental markets is sparse. Within the existing work, the focus remains centered on rural and/or amenity-based STRs such as beach-front and ski-resort access properties.  This focus results from the fact that, until a few years ago, the STR market itself operated almost wholly within these market segments. The shift to small, urban dwellings within the STR market is relatively new and explains the dearth of past work in this direction. With the rapid growth of Airbnb, and the controversies surrounding it, interest in the short-term rental market has increased among research as well as investors and industry analysts.

@@

Focusing specifically on the short-term rental market, @cassidy2007tourist suggest that within Australia's Gold Coast the STR industry lacks sophistication and efficiency.  Most notably, they find that most nightly or weekly rates are set by resident unit managers (RUMs) and are often under-priced as there is little accountability to owners.  The authors also note that very little research has been put toward understanding this growing market segment.

@@

Though issues surrounding second homes have been primarily treated as a rural phenomenon for decades  [@gallent2007second], a wave of ``new urban tourism'' [@fuller2014stop] has focused some of the problems of short-term rentals on large, urban environments. With this move to urban areas, issues such as nuisance, property rights and regulation have naturally arisen.  As a result, the short-term rental market has garnered significant interest in the public policy and legal sphere as of late. As many second homes are not permitted for or regulated as commercial entities, STRs can have serious effects on immediate neighbors in the way of externalies such as noise and congestion [@frost2006commentary].  While strict regulation is often proposed, doing so may constititute regulatory takings [@jefferson2015can].  Other, rights-based, regulatory approaches have been suggested [@miller2014transferable].  Many of the key legal decisions regarding short-term rentals, particulary through Airbnb, are still in progress and much remains to be settled in this area in terms of the regulatory and legal environment in which STRs are allowed to operate.

@@

Australia implements a particular tax policy, often termed `negative gearing', that allows for an income tax deduction to real estate investors who experience greater costs than revenues on investment properties.^[See @fane2004negative for a complete explanation of negative gearing.] To date, this policy only applies to long-term rental properties.  While this policy can influence the net revenue (post tax) of a property, the negative gearing policy requires knowledge of the investor's personal income situation.  As we do not have access to this information in the marketplace, we will not be considering any effects of this policy in our analysis here.

One benefit of a short-term rental may be the ability to adjust the nightly rate over the course of the year to reflect changing demand and market trends. As future market trends are unknown at the time of purchase and, within the span of 12 months, the decision of long- vs short-term leasing strategy is asymetrically irreversable (can go from short to long, but not long to short), we make our comparisons using a fixed nightly (short-term) rate to respresent the most likely state of knowledge for our hypothetical investor on September 1, 2015. Both long-term rents and short-term rates will be based on the market conditions as of this date.

@@

Janitorial costs, $J$, cover the costs to 'turn-over' the property between occupants.  For a short-term leasing strategy, these costs can be experienced every day or two, depending on the minimum length of stay and the total number of occupancies over the course of the year. Our data for short-term rentals comes from the Airbnb platform.  In addition to the nightly rate that an occupant pays to the owner there is also a separate 'cleaning fee' that is attached, as is a booking fee to Airbnb.^[This fee could be considered analogous to the leasing fees $L$ that long-term rentals pay.] Some hosts add additional 'linen' fees to cover costs of laundering.  While it may be advantageous for hosts in competitive markets to underquote their cleaning fees to keep occupancies up, we have no evidence to suggest that owners are exhibiting such behavior. As a result of janitorial fees being a separate line item in the STR data (paid separately and in addition to the nightly rate) they can also be ignored in the comparative analysis below.

@@

\begin{equation}
U = (4.5 * B * 1.5) * O
\end{equation}

@@

Unfortunately, there are no common identifiers between the two datasets as the Airbnb data does not include addresses. Both sets of data do include latitude and longitude, but in a dense urban area with many multiple-family dwellings it is impossible to match apartments based on two-dimensional coordinates. As a result of this, we are unable to directly match observations from the two datasets. However, the number of properties that rented long-term for the 2014 to 2015 period and then switched to Airbnb for the 2015 to 2016 period are likely very small and this inability to directly match properties does not greatly impact our modeling strategy.

@@

Our data preparation primarily involved the filtering out of data errors, outlier values and other observations than contain data that do not fit our research question. Additionally, as our data are from two different sources, a number of fields require standardization so that comparisons between the two leasing strategy options -- long-term rental (LTR) vs short-term rental (STR) -- can be made.

We begin by filtering observations by time. Our hypothetical leasing strategy decision is that of an investor who purchases a property on September 1, 2015 and must decide on traditional long-term rental versus an Airbnb-type, short-term rental approach to generating income. To determine the likely LTR nightly rate for any given property, we will use the observed rental transactions from September 1, 2014 to August 31, 2015. Taking the Airbnb or STR approach allows the owner to change rates over time and is subjected to daily changes in the market (supply and demand factors). In this simulated example we will use Airbnb data from September 1, 2015 to August 31, 2016 to represent the actual Airbnb market conditions over the time period in question.

@@

After removing property level observations that are missing price data, we find that most long-term weekly rentals are greater than \$200 and less than \$1,000 per week while most median nightly rates are greater than \$50 and less than \$500 per night. We remove observations outside of these ranges.

@@

To better visualize the relative frequency of the submarkets in the two tenure types we have broken down the data by submarket by property type by tenure type. A number of interesting observations are illustrated here (Figure 4). First, the vast majority of Airbnb units are apartments in city-core or city locations. As expected and as shown by the map (Figure 2), the long-term properties are more evenly spread across the metro region than the short-term properties.  Long-term houses are more peripheral than apartments, a result of urban development patterns and land economics. Beach properties are, relatively speaking, over-represented in the short-term market, especially in apartments, while suburban and rural apartments, somewhat common in long-term markets, are nearly absent in the short-term market. Overall, the suburban and rural Airbnb markets are quite thin, later results focusing on these markets will need to be examined in light of this small sample size.

@@

@@

First, we must deal with those properties that were not listed on Airbnb for the entire 12-month period.  Our data cleaning eliminated all properties that were listed for less than six months, however, it would be incorrect to compare the revenue of a property listed on Airbnb for seven months with the hypothetical annual revenue from a long-term rental.

@@

1. Actual ($REV_a$)}: Observed bookings multiplied by median nightly rate
2. Likely ($REV_l$)}: Observed bookings extrapolated to a full year representing the likely actual revenue if the property had been listed all year long. The additional bookings are imputed by comparing the property specific performance during the listed days to overall market performance and then applying that relative ratio to the overall market performance on the non-list days.  
3. Potential ($REV_p$)}: Observed bookings plus imputed bookings during blocked days.  The imputed bookings are imputed by comparing the property specific performance during the non-blocked days to overall market performance and then applying that relative ratio to the overall market performance on the blocked days.

As the 'Actual' calculation systematically under counts the revenue from properties that were not on the Airbnb portal for the entire year we disregard it going forward and consider the 'Likely' revenue metric to be a better estimate of market performance. We chart the distribution of the Likely ($REV_l$) and Potential ($REV_p$) short-term revenues against a distribution of the observed long-term revenues for the long-term properties (Figure 5). Revenues from the long-term rentals are far more consistent, with the vast majority totaling between \$12,000 to \$30,000 per year.  Likely short-term revenues range from neglible to highs of greater than \$80,000, with many of the properties showing very limited (less than \$10,000) revenue. The range of the distribution for the Potential short-term revenue is similar, however, the distribution itself is much flatter, with many more properties showing potential incomes in the \$20,000 to \$40,000 range.

While this simple analysis does not control for location and product differences, it does highlight the fact that some Airbnb owners are likely making greater revenues than in long-term rentals, but that most are likely not. The high number of very low yield Airbnb properties in the Likely condition suggests that some owners may not be focused on maximizing revenue, but instead prefer the flexibility of short-term leasing.^[Do note that we have not considered any overall market demand issues when analyzing potential revenue.]

![Revenue Comparison between Tenure Types](C:/Code/research/airbnbMelbourne/images/rev_dens.png)

@@

Realizing that the imputed long-term rental values may be more accurate when estimated at smaller scales, we impute long-term rental rates at submarket level, specifically by type and geographic area. In other words, we separately estimate the LTR model in Equation 10 for Apartments in the city core, Apartments within the beach submarket, etc. for all combinations of type and geographic area.

@@

where $GPR$ is the total number of guests allowed per bedroom^[Hosts set the total number of guests.  We have normalized this by bedroom to avoid co-linearity in our model with the bedroom variables.], $MinStay$ is the minimum number of nights that a stay may have and $CancelPolicy$ is the difficulty with which a booking may be cancelled.  We have grouped the cancellation policies into Flexible (24 hour notice), and Strict (those requiring 5 or more days notice, often with less than 100% refunds).

The results of all three models are shown in Table 4. For $PType$ Apartments are the reference category,  for $BedBath$, 1 Bed / 1Bath is the reference category, for $Submrkt$ the City-Core is the reference category and for $CancelPolicy$ Strict is the reference category.  As the additional variables are added, the model fit diagnostics steadily increase, suggesting that the increased complexity of the models adds to the quality of the fit.

In Model 1, we see that when controlling for structural characteristics, Houses have a significantly lower probability of being more profitable in short-term leasing than apartments. Additionally, 1 bed / 1 bath units show greater likelihood of short-term preference than 2 bed / 2 bath combinations. This could be due to the fact that 2 bed units may be very attractive in the long-term market as they are a useful product for two, non-family roomates to share. Larger units, 3 bedrooms plus, generally show no statistical difference in preference porbability to the 1 bed / 1 bath reference unit, with the exception of the 4 bed / 2 bath properties.

Adding the submarket variables (Model 2) shows that all areas have a much lower probability of short-term prefernce when compared to the City Core (the reference category) when controlling for structural characteristics.  The difference between the Rural and City-Core is not significant, however. The coefficients for the structural characteristics show the same signs and relatively small changes in magnitude from Model 1 to Model 2, speaking to the robustness of those estimates. The coefficient for House properties does decrease significantly in magnitude when adding the geographic submarkets.  This is likely due to the fact that there are sturctural differences between the geographic submarkets.  Houses, for example, are concentrated more heavily in City and Suburban areas than in the City-Core and some of the impact of House is being mediated by the geographic submarkets.

Finally, in Model 3 we add in host behavior or policies. Allowing a greater number of guests per bedroom results in a higher probability of short-term preference.  This variable may be operating as proxy for additional sleeping spaces which allow for a higher nightly rate, ceteris paribus, to be charged.  The greater the minimum stay, the lower the probability of short-term preference.  Longer minimum stays may equate to lower occupancy rates as one or two night travelers are exempted from the property.  Owners with longer minimum stays may be looking to minimize costs such as cleaning and administration.  Flexible cancellation policies result in a lower probability of short-term preference, while a Strict policy has a positive relationship with short-term preference. Strict cancellation may be a luxury for the best and highly desired properties on the market and, within our model, may be acting as a proxy for some unobserved measure of quality or demand while a flexible cancellation may be signalling the opposite.

@@

  # XTable
  x.full <- xtable(full.table,
                  caption='Logistic Modeling Results (Likely Revenue)\n')
  align(x.full) <- 'lllll'

@@

The above uses the Likely^[Extrapolated out in cases of Airbnb properties on the market for less than a year] revenue in the comparisons with the imputed long-term revenue.  Doing so may underestimate the possible revenue -- what may matter most to the hypothetical buyer in our scenario -- due to not factoring in the impact of blocked days.  If we, instead, assume that blocked days could be leased out on the short-term market at the same overall occupancy rate (relative for each property and the market as a whole on those days) as the unblocked (observed) days we can produce a potential revenue figure. Note that it is possible (likely?) that the large oversupply of available units in the market that would occur if no properties blocked their units would decrease both prices and occupancy rates.  We do not attempt to correct for this likelihood, rather we offer these figures as high end estimates.

Using the potential revenue, 36.1% of all properties are short-term preferable.  Breaking that down by type and by geographic submarket (Table 3) shows similar findings, in terms of relative comparativeness, to the Likely revenue analysis above.  Apartments have a higher short-term preference than do Houses and City-Core properties dominate the other geographic regions.  One notable difference here is that the Suburban and Rural properties perform rather poorly under a potential revenue scenario, relatively speaking, suggesting that they may be operating near their peak in the current condition.

@@

 px.full <- xtable(pfull.table,
                   caption='Logistic Modeling Results (Potential Revenue)\n')
 align(px.full) <- 'lllll'

@@

Overall, the choice of revenue type, Likely or Potential, does affect some of the conclusions to be derived from the results above.  As expected, under the Potential case there are more properties (36\% vs 28\%) that exhibit short-term preference.  Under the scenario, the possibility of increasing revenue by moving from long- to short-term leasing appears much better.  Again, however, we have not attempted to measure or estimate how this hypothetical increase of properties in the market (due to the elimation of blocked days) would impact overall market dynamics.  Suffice it to say that it would like drive down occupancies and maybe price, therefore leading to a decrease in short-term preference likelihood.

@@

If these are the dominant two forms of hosts in the market, then a distribution of properties by the percent of time which they are blocked should be highly bi-modal with profit seekers on the left and opportunistic sharers on the right. The left hand panel of Figure 6 shows that this breakdown holds somewhat, with peaks of hosts at each end, but that there are also a considerable number of hosts who block their properties some (25\% to 75\%) of the time.

There is also a third potential host type, that of the host who utilizies multiple platforms to list their short-term property, of which Airbnb may just be one. A potential method to identify these hosts is to look at the number of individual periods of blocked days.  A host using multiple platforms will likely have many small periods of blocked days on Airbnb as they will block out those days in which reservations are made on other platforms. We test to see if this is a meaningful metric by plotting the total count of blocked periods versus the percentage of blocked days in the right panel of Figure 6. We do find many hosts with a large number of blocked periods, but, unfortunately, no true pattern emerges.

@@

Attempts to divide the hosts up with unsupervised^[As there are no official host types we could not run more reliable supervised models.] clustering algorithms -- including the use of more variables -- failed to produce robust categories of hosts. As a more subjective, but implementable, approach we have divided the hosts up into four categories based on the original two variables considered: block percentage and number of blocked periods. Profit seekers are those host who have blocked percentages less than 25\% of the time.  Opportunistic Sharers have blocked their properties at least 75\% of the time.  Within those two extremes, we consider users who have at least 12 blocked periods during a year to be 'Multi-Platform Users' and the remaining set to be 'Unknown'. Profit Seekers make up 43\% of the hosts, with Opportunistic Sharers capturing 18\% and Multi-Platform Users 10\%.  The remainder 'Unknown' category contains just over 28\% of all properties. See Figures 7 and 8 for a breakdown of these group by the two variables.

@@

Looking at the differences in short-term preference under both revenue cases, we see large variation in short-term preference between host types in the Likely case and a slightly smaller variation in the Potential case, as expected. There is little change for the 'Profit Seekers' as most of these hosts are already operating with few blocked days.  For the other three sets, considerable gains could be had with a more aggresive leasing approach. As the potential figures are somewhat similar here, we will only examine the differences for host types in the Likely number going forward.

@@

Table 6 compares logistic models results from three of the host types.  The Opportunistic Sharers show very few instances of short-term preference and therefore could not be estimated.  The left hand column, that of the Profit-Seekers, is the most representative of a sample of property owners that are seeking to maximize revenue.  Overall, the coefficients from this sub-sample match up well to the results from the entire sample shown in Table 4.  In short, the conclusions are the same: small apartments in the central city with short minimal stays, inflexible cancelation policies and a high guest limit offer the highest probability of short-term preference.  The models for the Multi-Platform users (middle panel) and the Unknowns (right) are not as clear as the Profit-Seekers but do suggest overall similar dynamics.

@@

 xh.full <- xtable(hfull.table,
                  caption='Logistic Modeling Results (Potential Revenue by Host Type)\n')
 align(xh.full) <- 'lllll'

@@

In short, about 28\% of current Airbnb properties are outperforming their potential in the long-term market.  If owners all operated as profit maximizers this figure could be 36\% or higher given that the overall demand for short-term rental could match the assumed supply increase in such a case.
